An Examination of Sport Climbing Competition Format and Scoring System

### 2021-08-06 manuscript/Rmd/draft.Rmd

>>>>

Olympic Finalist Bassa Mawem pulled out of the sport climbing finals due to an injury in his left arm. Reuters reported that Megos, originally 9th in the total qualification rankings would not be replacing him. We suspect that this might be part of the unclear format of the Olympic Sport Climbing format. However, according to the climber's rankings, Megos technically should not have been next in line to fill in the missing finalist spot. Upon removing B. Mawem from the list and reranking the remaining competitors, we find that Chon would have moved into 8th place with a 640 combined score and Megos would have remained in 9th with 648. B. Mawem originally ranked first in the speed climbing, and Megos did comparatively poorly with 19th place. Both performed better than Mawem in bouldering and lead, leaving those rankings unchanged.

Prior, we considered a change in finalists if a non-finalist was removed. It is clear now though that the more interesting scenario is would would replace an original finalist who is removed from the rankings. Should it be the original 9th place competitor, or does it make more sense with such a scoring system to rerank? Surely Chon should be rewarded for his superior performance in speed compared to Megos with the chance to win a medal?

The next set of simulations will evaluate the probability that a removed finalist is replaced with someone other than the original 9th place competitor.

```{r, echo = FALSE}
load("draft_files/nineth_to_finals.rdata")

nineth_to_finals %>%
  ggplot(aes(x = dropped_rank, y = nineth_to_fin)) +
  geom_boxplot(aes(group = dropped_rank)) +
  theme_minimal() +
  xlab("Dropped Finalist Rank") +
  ylab("Probability that 9th Place Moves to Finals")
```
